Synthetic Observations of Extremely High Velocity (EHV) SiO component in young molecular outflows

### 2013-10-28 PAPER/Paola_1/ms_ver2.tex
Added the Physical Model in the paper

\newcommand{\rj}{R_{\rm jet}}

@@

Even with a myriad of such empirical evidences, the exact nature of
moleular emission in outflows is not clear. For a complete understanding it is imperative
to complement these observations with theoretical models. Many models
based on hydro-simulations and steady state shock calculations
were proposed to explain the observational signatures of molecular
outflows. Among them the two main models are that of wide-angled wind
driven \cite{Shu:1991p14071} and jet driven outflow
\cite{Canto:1991p14123}. 
The most popular among them is the jet driven model as wind
driven molecular outflows not only fail to match observed PV diagrams (\citealt{Cabrit:1992p14098})
but also tend to sweep large quantity of material at the extremities
of the lobes (\citealt{Masson:1992p14101}), while the jet driven models could successfully
derive the global outflow shapes and mass velocity relations of CO
outflows \citep{Raga:1993p9948,
  Masson:1993p9661,Downes:2003p9946,Downes:2007p9514}. 
There have also been some attempts to combine these two
models into one (\citealt{Shang:2006p14268}) to explain the global observational features. 
However, most of these dynamical models do not
account for shock chemistry. Instead, shock chemistry is studied
independently using steady state non-dissociative C-type and
dissociative J-type shocks models \citep{Neufeld:1989p11689, Schilke:1997p14140,Flower:2003p11236}. 
The MHD calculations by \cite{Glassgold:1991p13703} suggested that molecules
like SiO and CO could form within the jet. Similar conclusions of
molecules surviving in steady state disk winds have also been shown
(\citealt{Panoglou:2012p10039}). 
Very few simulations have modelled the outflow
dynamics including molecular chemistry and in absence of
magnetic fields \citep{Raga:1995p12965, Smith:2003p9985}. 
Recent models have also incorporated
infall from the envelope \citep{Rawlings:2013p14920} and radiation
transfer \citep{Offner:2011p14861} to explain molecular 
emission from Class 0 protostellar outflow sources.

@@

\subsection{Physical Model}
We model the propagation of jet as it interacts with the molecular
cloud core material at distances $\gtrsim$ 1000\,AU. 
At these distances from the central source, the downward pull of gravity plays a
negligible role and the dynamics of the jet is primarily governed by magnetic fields.
The total magnetic field in the jet is dominated by the toroidal
component. This is because the poloidal field decays as $z^{-2}$ as
compared to $z^{-1}$ for toroidal field to maintain the force balance
($z$ is the vertical distance from the source). 
The hoop stress due to pinch force from the toroidal magnetic fields maintains
a highly collimated beam like structure for the jet.
%

The ambient medium with which the jet interacts primarily
represents the molecular cloud core. The numerical domain is axi-symmetric
and in (r,z,$\phi$) cylindrical co-ordinates. Its extent 20 $R_{\rm j}$ in the radial
direction and 100
$R_{\rm j}$ along the vertical axis ($R_{\rm j}$ being the radius of
the jet). The domain of interest is about 3000\,AU away from the
central source and extends up to 0.1\,pc. Numerically, it is resolved by an uniform grid with a domain of 200$\times$1000 cells in the radial and vertical directions. For simplicity, we choose this medium
to be unmagnetized and non-turbulent. The density in the ambient
medium varies with vertical height $z$ as, $\rho_{\rm amb}\sim (\rho_0/z^{2})$
which is consistent with observations \citep{Keto:2010p15549, Caselli:2011p13935}. The value of
$\rho_0$ depends upon the density contrast, $\eta$, between the jet and
the ambient medium. The number density in the jet is kept fixed such
that the density in the ambient medium lies within a range of $10^{4}-10^{5}
\rm cm^{-3}$. The pressure in the ambient molecular medium is
set so to maintain a constant temperature of 50\,K. 
%

The jet enters into the medium through a nozzle of radius $R_{\rm jet}$
from the lower boundary (z = 0). The jet density is fixed to
be $10^{5} \rm cm^{-3}$ and it has a radius of $2.5\times10^{15} \rm cm \sim
167\,AU$. The jet is injected into the domain with a typical 
velocity of $v_{\rm jet,0}$ = 100 km/s as is the case for most low mass stellar jets
especially the low velocity component. The constant jet velocity is
superimposed with periodic pulsation of the form,
\begin{equation}
v_{\rm jet} = v_{\rm jet,0}(1.0 + A\,sin(2.0\pi t/T_{\rm p})),
\end{equation}
where the amplitude A is 0.25 and time period $T_{\rm p} = 70$
years. The pressure at the surface of the jet is $10^{-10} \rm
dyne\,\,cm^{-2}$ corresponding to a temperature of $T_{\rm jet} \sim
4\times10^{3}\,K$. Inside the jet beam, a radial variation of thermal
and magnetic pressure is adopted to maintain a magneto-static
equilibrium. We adopt the
same radial profiles used by \cite{Stone:2000p2650} for all our runs. 
Based on these profiles, the toroidal magnetic field is assumed to be
zero at the axis and achieves a maximum at some radius, $r_{\rm m}$
inside the jet. The maximum value, $B_{\phi,\rm m}$, depends on the
plasma $\beta$ which is kept fixed for all our runs to be at a value
of 10.0. This corresponds to the maximum field strength, $B_{\phi,\rm m}$ $\sim
16\mu\,G$.

@@

@@

\section{Chemistry and Cooling}
\label{sec:chem}
\subsection{Molecular Cooling}
\label{ssec:molcool}
Typically, early collimated outflows have cooling times shorter than
the dynamical times, implying that the radiative cooling plays a vital
role in dynamics. In order to study such outflows it is necessary to evolve
the dynamics in conjunction with full molecular chemistry and
associated radiative processes. However, evolving a full complex
network of chemical equations along with dynamics is a very
computationally intensive task. As a result, we have focused only in
evolving the chemical equations pertaining to atomic and molecular
hydrogen. This prescription of molecular cooling has been added to the PLUTO
code to study the chemical evolution of molecular outflows during the
very early phases of star formation. In particular, the total hydrogen
number density n$_{\rm H}$ comprises of contribution from atomic and
molecular hydrogen i.e., n$_{\rm H}$ = n$_{HI}$ + n$_{H2}$. Whereas
the contribution to electrons comes from the ionized hydrogen, n$_{\rm HII}$ and
from negligible but fixed fraction of metals ($Z \sim 10^{-4}$).
The chemical evolution of molecular, atomic and ionized hydrogen is governed by
equations listed in Table~\ref{tab:chemeq}. 
%

Numerically, the
non-homogenous parts of the equations~\ref{encons1} and \ref{chemevol} 
are are solved separately from the advection step
using operator splitting. Further, the rate equations are evolved at each advection time
step using sub-stepping and adaptive Runge-Kutta methods.
In doing so, the code tracks the formation and destruction of
fractions of three species viz., X(HI), X($H_{2}$)
and X(HII) based on the temperature dependent reaction rates specified
in the table along with their references and shown in the left panel
of figure~\ref{ssec:molcool}. The figure clearly shows that below
1000\,K the formation of molecular hydrogen dominates over other
chemical processes, whereas for hotter gas there exists a competition
of different chemical processes elucidating the complex chemical interplay.
The fractions of various hydrogen species are then used to update the cooling
function $\Lambda(\rho,T,{\bf{X}})$ to consistently derive the
temperature for next advection step. For low temperatures, we also
take into account contribution to cooling rates from molecules like
CO, whose abundance with respect to $n_{\rm H2}$ is fixed to
10$^{-5}$. For each of the other molecules contributing to the cooling function (like OH, H$_{2}$O and atomic OI)
the abundance is fixed to $5\times10^{-6}$. The contribution from
these molecules is only important when temperature is $<$ 100\,K as seen
in the right panel of figure~\ref{ssec:molcool}.

\begin{figure*}
 \includegraphics[width=1\columnwidth]{\figpath/H2ReactionRates_paper.png}
 \includegraphics[width=1\columnwidth]{\figpath/H2CoolingFuncs_full.png}
 \caption{Variation of $H_2$ chemistry reaction rates, $k_{i}$ and cooling
   function $\Lambda(n,T,{\bf{X}})$ with temperature for the initial
   state (see Sect.~\ref{ssec:molcool})}
\label{fig:tempvar}
\end{figure*}

@@

\subsection{Additional Cooling Prescriptions}
For comparison, we also ran simulations of axisymmetric MHD jets
with optically thin cooling prescriptions already existing in the public
release of the PLUTO-4 code. They are power-law cooling, 
simplified non-equilibrium (or atomic) cooling and tabulated
cooling. Further we compare these radiative runs with an adiabatic jet 
to cover a wide spectrum of jet thermodynamics. 
%

In case of power law cooling, the radiative losses denoted by
$\Lambda(\rho,T,{\bf X})$ are proportional to
$\rho^{2}T^{\alpha}$, with $\alpha$ = 0.5 representing bremsstrahlung
cooling. The loss term is independent of the fraction of various
species ${\bf X}$, therefore the fluid density $\rho$ is assumed to be
all hydrogen for further processing of emission maps. Similarly, 
for tabulated cooling prescription, the radiative losses are estimated
based on a fixed fraction of various species. However, the contribution
for each species is a complex function of density and temperature. The
variation of cooling function $\Lambda$ with temperature is shown in
the right panel of figure~\ref{fig:tempvar}. 
The formulations which depend on the fractions of each species and gas
temperature of these variations are mainly adopted from the appendix of \cite{Smith:2003p9985}. 
The fractions of all hydrogen species are kept fixed to their initial values, while that
of CO, OH, H$_{2}$O and OI are same as that used for molecular cooling
described above. The total cooling curve shown as the black solid line
in the figure~\ref{fig:tempvar} is used for the case of tabulated
cooling. For the atomic cooling, we distinguish between neutral
and ionized gas by defining $f_{\rm n} = n_{\rm HI}/n_{\rm H}$ and
evolve it using,
\begin{equation}
\frac{\partial f_{\rm n}}{\partial t} + {\bf{v}}\cdot\nabla f_{\rm n} =
n_{\rm e} [-(c_{\rm r} + c_{\rm i})f_{\rm n} + c_{\rm r}]
\end{equation}
where, $n_{\rm e}$ is the number of electrons that come from ionized
hydrogen and a very small fixed fraction of metals, c$_{\rm r}$ and c$_{\rm i}$ are the recombination and ionization rate coefficients that depend on the temperature T. In
addition to losses by collisional ionization and radiative
recombination, additional contribution to radiative losses come from various atomic line emissions from C,N,O,Mg,Si,Fe etc.

While these additional cooling prescriptions may not be a complete description of
the chemical and radiative processes in early outflows, they are ideal
to compare the results obtained from the more consistent newly added molecular
cooling. Such a comparison is presented in section~\ref{ssec:coolres}.

@@

Further, the velocities obtained from the dynamical simulations are in fact the
jet flow velocities. They can well be different from the shock
velocities depending upon the density contrast between the jet and
ambient medium. For the present study, we will consider two
cases. Firstly, the most simple one in which the jet flow velocity is
same as the shock velocity. Considering that there is a density
contrast in the simulation setup, this assumption is will
overestimate the shock velocity. We also
consider a case in which the shock speed of a dense plug accumulated
between the jet and medium is estimated by balancing the respective ram
pressures \citep{Masson:1993p9661},
\begin{equation}
\label{eq:shockjetvel}
v_{\rm s} = \frac{v_{\rm j}}{1 + \eta^{-0.5}} = v_{\rm j}\delta
\end{equation}
where v$_{\rm s}$ is the velocity of dense plug and v$_{\rm j}$ is the
jet velocity. A comparative view of using such velocity corrections and
functional forms of SiO abundances are discussed in
Sec~\ref{ssec:sioabunem}.

@@

A value of $\eta > 1$, means that
that jet is over dense with respect to the medium. In all our runs with
different cooling prescriptions, we have assumed the jet to be
over-dense by 10 times compared to the ambient medium. Additionally, for
the atomic, tabulated and molecular cooling runs (see section~\ref{sec:chem}
for description of different cooling prescriptions) we have also used a
value of $\eta$ of the order of unity indicating similar densities in
the jet and the medium. 
The magnetic field strength is kept to be fixed using $\beta = 10.0$,
for all our runs. Table~\ref{tab:result1} lists all the runs with
varying $\eta$ and cooling prescriptions, along with the peak
intensity and line widths at the bow shock.
%

@@

\begin{table*}
\centering
\caption{Summary from parameter runs. The parameter $\delta$, is the ratio of
  shock to jet velocity and it depends on the density contrast $\eta$
  as shown in eq.~\ref{eq:shockjetvel}}
\begin{tabular}{c | c | c | c | c | c | c }
\hline
Run & Cooling Mode & $\eta$ & \multicolumn{2}{|c|}{Top Hat Profile
  $\delta = 1$} & \multicolumn{2}{|c|}{Gaussian Profile $\delta =
  1$}\\
\hline\hline
&&& $\int T_{\rm MB}dV$ [K-km\,s$^{-1}$] & $\Delta$v [km s$^{-1}$] & $\int T_{\rm MB}dV$ [K-km\,s$^{-1}$] & $\Delta$v [km s$^{-1}$] \\
\hline
adi1010 & Nil (Adiabatic) & 10 & 2.91 & $>$40 & 0.25 & 40.0 \\
pow1010 & Power law & 10 & 0.64 & 8.0 & 0.02 & 10.0\\
atm1010 & Atomic & 10 & 2.04 & 36.0 & 0.58 & 38.0 \\
atm210 & Atomic & 2 & 3.21 & 18.0 & 0.64 & 20.0 \\
tab1010 & Tabulated & 10 & 0.75 & 11.0 & 0.09 & 10.0 \\
tab210 & Tabulated & 2 & 2.89 & 8.0 & 1.0 & 9.0 \\
mol1010 & Molecular & 10 & 1.10 & 10.0 & 0.1 & 11.0 \\
mol310 & Molecular & 3 & 3.3 & 14.0 & 0.66 & 12.0\\
\hline
\end{tabular}
\label{tab:result1}
\end{table*}

@@

\subsection{Reference Run}
\label{ssec:refrun}
We define a reference run in order to quantify and compare results obtained from such a
survey of above mentioned parameters. The results shown in this work will be
pertaining to the reference run (unless otherwise mentioned) and appropriate comparison will be
discussed with other runs.
%

The reference run in our calculation has density contrast $\eta$ = 3
with a plasma beta $\beta$ = 10. The jet for this run is 89\% atomic,
10\% molecular and 1\% ionized to begin with. This jet enters the
ambient medium with a velocity of v$_{\rm jet,0}$ = 100 km
s$^{-1}$. The cooling in the jet is
via molecular cooling prescription, whereby the hydrogen fractions are
evolved with cooling contributions for abundant molecules like CO, OH
etc. The final state (at time $\sim$ 1000 yr), i.e, density, temperature and velocity obtained from this 
run are given in input to the radiative transfer code to obtain the observational features
corresponding to SiO molecule. 
For the radiative transfer calculation, the source in the reference
run is stationary and the jet is in the plane of the sky ($\phi =
90^{\circ}$). A gaussian profile is used for the fractional abundance
of SiO.
%TheSiO fractional abundance used is that obtained
%from equation~\ref{eq:fracSio}. 
Further, the shock velocity which determines
the fractional abundance, v$_{\rm s}$, is less than the jet velocity in the
flow, such that their ratio $\delta$ is given by
equation~\ref{eq:shockjetvel}, where $\eta$ is the density contrast as a
function of height from the base of the jet.

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/diffcool_sim.pdf}%{\figpath/jetplt1_1010.png}
 \caption{Jet Volume Density for different cooling modes with
   $\eta$ = 10 and $\beta$ = 10.}
\label{fig:coolcmpdyn}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/diffcool_img.pdf}
 \caption{A plot of the integrated SiO J=2-$>$1 emission from 5 models, 
each using a different method to calculate cooling and all with $\eta$=10 $\beta$=10.}
\label{fig:coolcmpsio} 
\end{figure*}

% \subsection{Effects of changing $\eta$ and $\beta$}
% figure nos. 2 -- Simulated images describing lengths, collimation and
% emission for lower eta more stronger. 
% \begin{figure*}
%  \includegraphics[width=2.\columnwidth]{\figpath/jetplt2_diffetabeta.png}
%  \caption{Jet Volume Density for atomic cooling with $\eta$ = 10 but $\beta$ = 1 and
%    10 and that with tabulated cooling with $\beta$ = 10 but $\eta$ = 2
%    and 10.}
% \label{grid}
% \end{figure*}

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/diffabun_img.pdf}%{\figpath/EM_gaussandth.png}
 \caption{Variation of 2-$>$1 SiO emission for runs with molecular
   cooling having $\eta$ = 3 and $\beta$ = 10 and different abundance profiles.}
\label{fig:diffabunem}
\end{figure*}

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/diffabun_PV.pdf}%{\figpath/PV_gaussandth.png}
 \caption{Position-velocity maps of SiO(2-1) for the internal
   knots produced in the model with molecular
   cooling, $\eta$ = 3, $\beta$ = 10 and different abundance
   profiles. The contours mark different levels of emission in Kelvins, viz.,
   0.2,0.6,1.0,1.4,1.8,2.0,3.0,4.0.}
\label{fig:pvdiffabun}
\end{figure*}

>>>>

\newcommand{\rj}{R$_{\rm jet}$}

@@

Even with a myriad of such empirical evidences, the exact nature of
molecular emission in outflows is not clear. For a complete understanding it is imperative
to complement these observations with theoretical models. Many models
based on hydro-simulations and steady state shock calculations
were proposed to explain the observational signatures of molecular
outflows. Among them the two main models are that of wide-angled wind
driven \cite{Shu:1991p14071} and jet driven outflow
\cite{Canto:1991p14123}. 
The most popular among them is the jet driven model as wind
driven molecular outflows not only fail to match observed PV diagrams (\citealt{Cabrit:1992p14098})
but also tend to sweep large quantity of material at the extremities
of the lobes (\citealt{Masson:1992p14101}), while the jet driven models could successfully
derive the global outflow shapes and mass velocity relations of CO
outflows \citep{Raga:1993p9948,
  Masson:1993p9661,Downes:2003p9946,Downes:2007p9514}. 
There have also been some attempts to combine these two
models into one (\citealt{Shang:2006p14268}) to explain the global observational features. 
However, most of these dynamical models do not
account for shock chemistry. Instead, shock chemistry is studied
independently using steady state non-dissociative C-type and
dissociative J-type shocks models \citep{Neufeld:1989p11689, Schilke:1997p14140,Flower:2003p11236}. 
The MHD calculations by \cite{Glassgold:1991p13703} suggested that molecules
like SiO and CO could form within the jet. Similar conclusions of
molecules surviving in steady state disk winds have also been shown
(\citealt{Panoglou:2012p10039}). 
Very few simulations have modelled the outflow
dynamics including molecular chemistry and in absence of
magnetic fields \citep{Raga:1995p12965, Smith:2003p9985}. 
Recent models have also incorporated
infall from the envelope \citep{Rawlings:2013p14920} and radiation
transfer \citep{Offner:2011p14861} to explain molecular 
emission from Class 0 protostellar outflow sources.

@@

\subsection{Physical Model}
We model the propagation of an axi-symmetric jet
in (r,z,$\phi$) cylindrical co-ordinates as it interacts with the
ambient medium at distances $\gtrsim$ 1000\,AU. 
At these distances from the central source, the downward pull of gravity plays a
negligible role on the dynamics of the jet. The ambient medium with which the jet interacts primarily
represents the molecular cloud core. We study the interaction within a
radial extent of 20\,\rj\,\, and a vertical extent of 100\,\rj\,\,(\rj\,\, being the radius of
the jet). Numerically, it is resolved by an uniform grid with a domain
of 200$\times$1000 cells in the radial and vertical directions. For simplicity, we choose this medium
to be unmagnetized and non-turbulent. The density in the ambient
medium varies with vertical height $z$ as, $\rho_{\rm amb}\sim (\rho_0/z^{2})$
which is consistent with observations \citep{Keto:2010p15549,
  Caselli:2011p13935}. The value of
$\rho_0$ depends upon the density contrast, $\eta$, between the jet and
the ambient medium. The number density in the jet is kept fixed such
that the density in the ambient medium lies within a range of $10^{4}-10^{5}
\rm cm^{-3}$. The pressure in the ambient molecular medium is
set so to maintain a constant temperature of 50\,K. Note, dense
molecular cores through which young outflows propagate
have lower temperatures ranging from 10-20\,K
(\citealt{Rathborne:2007p15753}). As the jet is hypersonic in both
cases, runs with lower ambient temperature of 10\,K do not show any 
appreciable changes in final dynamical structure of the jet.

@@

The jet enters into the medium through a nozzle of radius \rj\,\,
from the lower boundary (z = 0). Optically visible jets from T-Tauri
stars are just about spatially resolved for the nearest
star forming regions i.e, their radii are of order
100\,AU at a few thousand AU from the source \citep{Ray:2012p15733}.
These jets have mass outflow rates of 10$^{-7}$ M$_{\odot}$
yr$^{-1}$ \citep{Dougados:2010p15774}. Younger and 
embedded jets in their Class 0 phase have
outflow rates about 10-100 times higher than 
corresponding jets from optically visible young stars
\citep{Bally:2007p2340, Dionatos:2009p15670, Ray:2012p15733}. Also,
these optically obscure jets are relatively slower with velocities
ranging around 100 km\,s$^{-1}$ \citep{Bally:2007p2340} as compared to
optically visible flows with velocities up to 500 km s$^{-1}$
\citep[e.g,][]{Mundt:1987p15799}.
For the present model, the jet is injected into the domain with a 
velocity of $v_{\rm jet,0}$ = 100 km/s and set the jet density to
$10^{5} \rm cm^{-3}$ within a radius of $2.5\times10^{15} \rm
cm$. With these choices we obtain a mass outflow rate of 10$^{-6}$
M$_{\odot}$ yr$^{-1}$. Most of the young outflows 
show variable features possibly due to variable accretion rates
or internal shocks inside the jet \cite{Bachiller:1996p4692}.
We replicate such variability by superimposing the constant jet
velocity with sinusoidal pulsation with an amplitude of 25
km\,s$^{-1}$ with a periodicity of 70\,yr such that the dense knots
are separated by a distance of $\sim$\,1000\,AU.

Thermal and magnetic properties of jets from young stars are difficult
to constraint because of their large variation not only within the jet
but also in its surroundings. Forbidden line emissions from
HH objects indicate excitation temperatures of the order
of $\leq$10$^{4}$\,K \citep{Podio:2006p15921,
  Bally:2007p2340}. However, in case of 
young jets the temperature is
considerably lower (i.e, by an order of magnitude) than found in jets from Class I/II
sources as derived from various atomic lines in molecular jets from
L-1448C and HH211 \citep{Dionatos:2009p15670,Dionatos:2010p15968}.  
The temperature at the jet
nozzle for our model has been set to 1000\,K. Thus, the jet is
over-pressured by 60 times as compared with the ambient medium,
assuming $\eta$ = 3.0. Inside the jet beam, a radial variation of thermal
and magnetic pressure is adopted to maintain a magneto-static
equilibrium for all our runs. 
Based on these profiles (adopted from \citealt{Stone:2000p2650}), 
the toroidal magnetic field is assumed to be
zero at the axis and achieves a maximum at radius, $r_{\rm m} \sim 0.8$\rj\,\,
inside the jet. Though measuring magnetic fields in
most jets is not possible via unpolarized nebular lines, in some special
cases like HH 34S and HH 111V, it has been measured to be 10 and 30
$\mu\,G$ respectively \citep{Morse:1993p15960}. Considering the
dependence of such weak magnetic fields on density and radial distance in velocity-variable flows, 
\cite{Hartigan:2007p15768} has shown that at large distances of the 
star these fields only play a minor role in the dynamics. Thus for all
our runs the value of magnetic field strength, $B_{\phi}(r_{\rm m})$ 
is set to be $\sim 37\mu\,G$.
%

@@

\section{Chemistry and Cooling}
\label{sec:chem}
\subsection{Molecular Chemistry}
\label{ssec:molcool}
Typically, early collimated outflows have cooling times shorter than
the dynamical times, implying that the radiative cooling plays a vital
role in dynamics. In order to study such outflows it is necessary to evolve
the dynamics in conjunction with full molecular chemistry and
associated radiative processes. However, evolving a full complex
network of chemical equations along with dynamics is a very
computationally intensive task. As a result, we have focused only in
evolving the chemical equations pertaining to atomic and molecular
hydrogen. This prescription of molecular cooling has been added to the PLUTO
code to study the chemical evolution of molecular outflows during the
very early phases of star formation. In particular, the total hydrogen
number density n$_{\rm H}$ comprises of contribution from atomic and
molecular hydrogen i.e., n$_{\rm H}$ = n$_{HI}$ + n$_{H2}$. Whereas
the contribution to electrons comes from the ionized hydrogen, n$_{\rm HII}$ and
from negligible but fixed fraction of metals ($Z \sim 10^{-4}$).
%
Numerically, the
non-homogenous parts of the equations~\ref{encons1} and \ref{chemevol} 
are are solved separately from the advection step
using operator splitting. Further, the rate equations are evolved at each advection time
step using sub-stepping and adaptive Runge-Kutta methods.
The formation and destruction of
fractions (denoted by \textbf{X}$_{i}$ in eq.~\ref{chemevol}) of three species, viz., fHI, f$H_{2}$
and fHII are tracked by the code based on the temperature dependent reaction rates
whose coefficients are specified 
in the Table~\ref{tab:chemeq} along with their literature references.

%and shown in the left panel
%of figure~\ref{fig:tempvar}. The figure clearly shows that below
%1000\,K, the rate coefficient of radiative recombination is the
%most dominant. However at these temperatures, very few
%electrons and ionized hydrogen are present in the initial ambient
%medium, thus the corresponding rate for this reaction is not that
%significant initially. At low temperatures corresponding to the molecular
%cores, the rate of formation of molecular hydrogen dominates over other
%chemical processes, whereas for gas temperatures $>$ 1000\,K there exists a competition
%of different chemical processes elucidating the complex chemical interplay.

\begin{figure}
% \includegraphics[width=1\columnwidth]{\figpath/H2ReactionRates_paper.png}
 \includegraphics[width=1\columnwidth]{\figpath/H2CoolingFuncs_full.png}
 \caption{Variation of the cooling
   function $\Lambda(n,T,{\bf{X}})$ with temperature for the initial
   state}
\label{fig:tempvar}
\end{figure}

@@

\subsection{Cooling Prescriptions}
%
Figure~\ref{fig:tempvar} shows the variation
of the cooling function, $\Lambda(\rho,T,{\bf{X}})$, with gas
temperature assuming conditions of initial ambient density. 
Various processes that contribute to the cooling function in our model
are listed in the figure. The formulations of these contributions
are adopted from \citep[][and references therein]{Smith:2003p9985, OSullivan:2009p10078}.
For low temperatures, we also 
take into account contribution to cooling function from molecules like
CO, whose abundance with respect to $n_{\rm H2}$ is fixed to
10$^{-5}$. For each of the other molecules contributing to the cooling function (like OH, H$_{2}$O and atomic OI)
the abundance is fixed to $5\times10^{-6}$. The contribution from
these molecules is only important when temperature is $<$ 100\,K as seen
in figure~\ref{fig:tempvar}. A substantial contribution to the cooling
function, between T = 100\,K and 3000\,K, 
comes from the vibrational and rotational transitions of
molecular hydrogen. Whereas, for temperatures above 3000\,K the
cooling is mainly via H$_{2}$ disassociation. The total curve for the
whole temperature range obtained in our runs is shown in black.
The fact that the initial ambient medium is largely molecular, cooling is also dominated by
molecular processes. In our reference model (see ~\ref{ssec:refrun}), 
we update this curve at each time step based on the fraction of
various hydrogen species. Thus during the evolution, even
contributions from collisional ionizations and excitations also become
important in regions dominated by atomic and ionized hydrogen.
%

We also ran simulations of axisymmetric MHD jets
with optically thin cooling prescriptions already existing in the public
release of the PLUTO-4 code. They are power-law cooling, 
simplified non-equilibrium (or atomic) cooling and tabulated
cooling. While these additional cooling prescriptions may not be a complete description of
the chemical and radiative processes in early outflows, they are ideal
to compare the results obtained from the more consistent newly added molecular
cooling. In addition, we compare these radiative runs with an adiabatic jet 
to cover a wide spectrum of jet thermodynamics. 
Such a comparison is presented in section~\ref{ssec:coolres}.

@@

Further, the velocities obtained from the dynamical simulations are in fact the
jet flow velocities. They can well be different from the shock
velocities depending upon the density contrast between the jet and
ambient medium. For the present study, we will consider two
cases. Firstly, the most simple one in which the jet flow velocity is
same as the shock velocity. Considering that there is a density
contrast in the simulation setup, this assumption is will
overestimate the shock velocity. We also
consider a case in which the shock speed of a dense plug accumulated
between the jet and medium is estimated by balancing the respective ram
pressures \citep{Masson:1993p9661},
\begin{equation}
\label{eq:shockjetvel}
v_{\rm s} = \frac{v_{\rm jet}}{1 + \eta^{-0.5}} = v_{\rm jet}\delta
\end{equation}
where v$_{\rm s}$ is the velocity of dense plug and v$_{\rm jet}$ is the
jet velocity. A comparative view of using such velocity corrections and
functional forms of SiO abundances are discussed in
Sec~\ref{ssec:sioabunem}.

@@

A value of $\eta > 1$, means that
that jet is over dense with respect to the medium. In all our runs with
different cooling prescriptions, we have assumed the jet to be
over-dense by 10 times compared to the ambient medium. Additionally, for
the atomic, tabulated and molecular cooling runs (see section~\ref{sec:chem}
for description of different cooling prescriptions) we have also used a
value of $\eta$ of the order of unity indicating similar densities in
the jet and the medium. Table~\ref{tab:result1} lists all the runs with
varying $\eta$ and cooling prescriptions, along with the peak
intensity and line widths at the bow shock.
%

@@

\begin{table*}
\centering
\caption{Summary from parameter runs. The parameter $\delta$, is the ratio of
  shock to jet velocity and it depends on the density contrast $\eta$
  as shown in eq.~\ref{eq:shockjetvel}}
\begin{tabular}{c | c | c | c | c | c | c }
\hline
Run & Cooling Mode & $\eta$ & \multicolumn{2}{|c|}{Top Hat Profile
  $\delta = 1$} & \multicolumn{2}{|c|}{Gaussian Profile $\delta =
  1$}\\
\hline\hline
&&& $\int T_{\rm MB}dV$ [K-km\,s$^{-1}$] & $\Delta$v [km s$^{-1}$] & $\int T_{\rm MB}dV$ [K-km\,s$^{-1}$] & $\Delta$v [km s$^{-1}$] \\
\hline
adi10 & Nil (Adiabatic) & 10 & 2.91 & $>$40 & 0.25 & 40.0 \\
pow10 & Power law & 10 & 0.64 & 8.0 & 0.02 & 10.0\\
atm10 & Atomic & 10 & 2.04 & 36.0 & 0.58 & 38.0 \\
atm2 & Atomic & 2 & 3.21 & 18.0 & 0.64 & 20.0 \\
tab10 & Tabulated & 10 & 0.75 & 11.0 & 0.09 & 10.0 \\
tab2 & Tabulated & 2 & 2.89 & 8.0 & 1.0 & 9.0 \\
mol10 & Molecular & 10 & 1.10 & 10.0 & 0.1 & 11.0 \\
mol3 & Molecular & 3 & 3.3 & 14.0 & 0.66 & 12.0\\
\hline
\end{tabular}
\label{tab:result1}
\end{table*}

@@

\subsection{Reference Run}
\label{ssec:refrun}
We define a reference run in order to quantify and compare results obtained from such a
survey of above mentioned parameters. The results shown in this work will be
pertaining to the reference run (unless otherwise mentioned) and appropriate comparison will be
discussed with other runs.

The reference run in our calculation has density contrast $\eta$ = 3
with the initial jet density of 10$^{5}$ cm$^{-3}$. This jet enters the
ambient medium with a velocity of v$_{\rm jet,0}$ = 100 km
s$^{-1}$ which is further superimposed by sinusoidal perturbations to
replicate dense knots. The maximum value of toroidal magnetic field is
set to be 37$\mu\,G$ at the base of the jet.
\cite{Dionatos:2009p15670} have detected atomic lines
towards a molecular jet from L1448-C indicating a presence of embedded
atomic jet at low excitation. Additionally, they also detect pure
rotational transitions from H$_{2}$ indicating presence of molecules
along with the atomic jet. Line ratios from this source indicate that the
atomic gas is characterized by an an ionization fraction of 1\%,
which is considerably less as compared to ionization fraction of
10-20\% measured in optical jets \cite{Bacciotti:1995p15970}.
Initial jet composition is prescribed
be 89\% atomic, 10\% molecular and 1\% ionized, in qualitative
agreement of above observational results. 
%

The cooling in the reference jet is via molecular cooling prescription, whereby
the initial hydrogen fractions are evolved using appropriate rate equations and with cooling 
contributions for abundant molecules like CO, OH
etc. The final state (at time $\sim$ 1000 yr), i.e, density, temperature and velocity obtained from this 
run are given as inputs to the radiative transfer code to obtain the observational features
corresponding to SiO molecule. 
For the radiative transfer calculation, the source in the reference
run is stationary and the jet is in the plane of the sky ($\phi =
90^{\circ}$). A gaussian profile without any upper cutoff is used for the fractional abundance
of SiO. Further, the shock velocity which determines
the fractional abundance, v$_{\rm s}$, is less than the jet velocity in the
flow, such that their ratio $\delta$ is given by
equation~\ref{eq:shockjetvel}, where $\eta$ is the density contrast as a
function of height from the base of the jet.

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/diffcool_sim.pdf}%{\figpath/jetplt1_1010.png}
 \caption{Jet Volume Density for different cooling modes with
   $\eta$ = 10.}
\label{fig:coolcmpdyn}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/diffcool_img.pdf}
 \caption{A plot of the integrated SiO J=2-$>$1 emission from 5 models, 
each using a different method to calculate cooling and all with $\eta$=10.}
\label{fig:coolcmpsio} 
\end{figure*}

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/diffabun_img.pdf}%{\figpath/EM_gaussandth.png}
 \caption{Variation of 2-$>$1 SiO emission for runs with molecular
   cooling having different abundance profiles.}
\label{fig:diffabunem}
\end{figure*}

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/diffabun_PV.pdf}%{\figpath/PV_gaussandth.png}
 \caption{Position-velocity maps of SiO(2-1) for the internal
   knots produced in the model with reference parameters and different abundance
   profiles. The contours mark different levels of emission in Kelvins, viz.,
   0.2,0.6,1.0,1.4,1.8,2.0,3.0,4.0.}
\label{fig:pvdiffabun}
\end{figure*}
